---
title: 'Example in Early Breast Cancer'
author: "Javier Sanchez Alvarez and Valerie Aponte Ribero"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    embed-resources: true
    self-contained-math: true
vignette: >
  %\VignetteIndexEntry{example_eBC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

# Introduction
This document runs a discrete event simulation model in the context of early breast cancer to show how the functions can be used to generate a model in only a few steps.

When running a DES, it's important to consider speed. Simulation based models can be computationally expensive, which means that using efficient coding can have a substantial impact on performance.

## Main options
```{r setup}
library(WARDEN)

library(purrr)
library(dplyr)
library(ggplot2)
library(kableExtra)

```


## Model Concept
Patients start in early breast cancer, and draw times to event. Patients also draw a probability of going into metastatic breast cancer or going into remission. If they go into remission, they can have a metastatic recurrence. At any point in time they can die, depending on the risk of each disease stage.


## Load Data
The dummy data for costs and utility is generated below.

```{r load_data}
#Utilities
df_util <- data.frame( name = c("util.idfs.ontx" ,"util.idfs.offtx" ,"util.remission" ,"util.recurrence" ,"util.mbc.progression.mbc" ,"util.mbc.pps"),
                         value = c(0.75, 0.8,0.9,0.7,0.6,0.5),
                         se=rep(0.02,6),
                         stringsAsFactors = FALSE
)


#Costs
df_cost <- data.frame( name = c("cost.idfs.tx" ,"cost.recurrence" ,"cost.mbc.tx" ,"cost.tx.beva" ,"cost.idfs.txnoint",
                                  "cost.idfs","cost.mbc.progression.mbc","cost.mbc.pps","cost.2ndline","cost.ae"),
                         value = c(40000,5000,3000,10000,30000,
                                   10000,20000,30000,20000,1000),
                         stringsAsFactors = FALSE
) %>%
  mutate(se= value/5)

```

# General inputs with delayed execution
Initial inputs and flags that will be used in the model can be defined below. We can define inputs that are common to all patients (`common_all_inputs`) within a simulation, inputs that are unique to a patient independently of the treatment (e.g. natural death, defined in `common_pt_inputs`), and inputs that are unique to that patient and that treatment (`unique_pt_inputs`). Items can be included through the `add_item` function, and can be used in subsequent items (e.g. below, we define `sex_pt` and we use it in `nat.os.s` to get the background mortality for that patient). All these inputs are generated before the events and the reaction to events are executed. Furthermore, the program first executes `common_all_inputs`, then `common_pt_inputs` and then `unique_pt_inputs`. So one could use the items generated in `common_all_inputs` in `unique_pt_inputs`.

The flag `fl.remission` is drawn using a Bernoulli distribution with probability 0.8. This means that 80% of the patients will have a remission, while 20% will go into early metastatic BC. Note that this could also be modeled differently by using a time to remission and time to early metastatic BC, comparing these and choosing the pathway depending on which one is smaller. 

We also define here the specific utilities and costs that will be used in the model. We create a named vector that identifies the costs and utilities, and then we assign them to each specific item. Note that while `util_v` or `cost_v` is a named vector, the specific items that will be used by the model (e.g. `util.remission`) are unnamed. It is strongly recommended to assign unnamed objects if they are going to be processed in the model. In this case, we're only using `util_v` and `cost_v` as an intermediate input and these objects will not be processed (we just use them to make the code more readable), so it's fine if we name them. However, naming the value of `util.remission` (e.g. using one bracket instead of two in `util_v[["util.remission"]]`) may cause the outputs from the model to change names, depending on use. This is because of how R works: it would correspond to a named list with a named vector/element, which R concatenates, so in this case it would end up generating `qaly.util.remission` as an output of the model instead of just `qaly`).

```{r input_delayed}
#Each patient is identified through "i"
#Items used in the model should be unnamed numeric/vectors! otherwise if they are processed by model it can lead to strangely named outcomes
#In this case, util_v is a named vector, but it's not processed by the model. We extract unnamed numerics from it.

#Put objects here that do not change on any patient or intervention loop
common_all_inputs <- add_item() %>%
  add_item( #utilities
  pick_val_v(
    base =  df_util$value,
    psa = MASS::mvrnorm(1,df_util$value,diag(df_util$se^2)),
    sens = df_util$value,
    psa_ind = psa_bool,
    sens_ind = sensitivity_bool,
    indicator = rep(0, nrow(df_util)),
    names_out =df_util$name
  ),#costs
  pick_val_v(
    base =  df_cost$value,
    psa = rgamma_mse(1,df_cost$value,df_cost$se),
    sens = df_cost$value,
    psa_ind = psa_bool,
    sens_ind = sensitivity_bool,
    indicator = rep(0, nrow(df_cost)),
    names_out =df_cost$name
  )
)
#Put objects here that do not change as we loop through interventions for a patient
common_pt_inputs <- add_item(sex_pt = ifelse(rbinom(1,1,p=0.01),"male","female"),
                             nat.os.s = rcond_gompertz(1,
                                                     shape=if(sex_pt=="male"){0.102}else{0.115},
                                                     rate=if(sex_pt=="male"){0.000016}else{0.0000041},
                                                     lower_bound = 50) ) #in years, for a patient who is 50yo

#Put objects here that change as we loop through treatments for each patient (e.g. events can affect fl.tx, but events do not affect nat.os.s)
#common across arm but changes per pt could be implemented here (if (arm==)... )
unique_pt_inputs <- add_item(
  fl.idfs.ontx             = 1,
  fl.idfs                  = 1,
  fl.mbcs.ontx             = 1,
  fl.mbcs.progression.mbc  = 1,
  fl.tx.beva               = 1,  
  fl.mbcs                  = 0,
  fl.mbcs_2ndline          = 0,
  fl.recurrence            = 0,
  fl.remission             = rbinom(1,1,0.8), #80% probability of going into remission
  q_default = if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              },
  c_default = if(arm=="noint"){cost.idfs.txnoint* fl.idfs.ontx  + cost.idfs}else{(cost.idfs.tx) * fl.idfs.ontx + cost.tx.beva * fl.tx.beva + cost.idfs},
  c_ae = 0
)
```

# Events
## Add Initial Events
Events are added below through the `add_tte` function. We use this function twice, one per intervention. We must define several arguments: one to indicate the intervention, one to define the names of the events used, one to define the names of other objects created that we would like to store (optional, maybe we generate an intermediate input which is not an event but that we want to save) and the actual input in which we generate the time to event. Events and other objects will be automatically initialized to `Inf`. We draw the times to event for the patients. This chunk is a bit more complex, so it's worth spending a bit of time explaining it.

The `init_event_list` object is populated by using the `add_tte` function twice, one for the "int" strategy and other for the "noint" strategy. We first declare the `start` time to be `0`.

We then proceed to generate the actual time to event. We use the `draw_tte()` function to generate the time to event using a log-normal distribution for the event variables that are of interest. One should always be aware of how the competing risks interact with each other. While we have abstracted from these type of corrections here, it is recommended to have an understanding about how these affect the results and have a look at the competing risks/semi-competing risks literature.

Note that in our model, the initial list of events are `start, ttot, ttot.beva, progression.mbc, os, idfs, ttot.early, remission, recurrence and start.early.mbc`. However, other, non-initial events can be defined in the reactions part seen in the section below.

```{r model_evts}
init_event_list <- 
  add_tte(arm="int",
               evts = c("start","ttot", "ttot.beva","progression.mbc", "os","idfs","ttot.early","remission","recurrence","start.early.mbc","ae","2ndline_mbc"),
               other_inp = c("os.early","os.mbc"),
               input={ #intervention
    start <- 0
    
    #Early
    
    idfs <- draw_tte(1,'lnorm',coef1=2, coef2=log(0.2)) 
    ttot.early <- min(draw_tte(1,'lnorm',coef1=2, coef2=log(0.2)),idfs)
    ttot.beva <- draw_tte(1,'lnorm',coef1=2, coef2=log(0.2))
    
    os.early <- draw_tte(1,'lnorm',coef1=3, coef2=log(0.2)) 
    
    #if patient has remission, check when will recurrence happen
    if (fl.remission) { 
      recurrence <- idfs +draw_tte(1,'lnorm',coef1=2, coef2=log(0.2))
      remission <- idfs
      
      #if recurrence happens before death
      if (min(os.early,nat.os.s)>recurrence) { 
        
        #Late metastatic (after finishing idfs and recurrence)
        
        os.mbc <- draw_tte(1,'lnorm',coef1=0.8, coef2=log(0.2)) + idfs  +  recurrence 
        
        progression.mbc <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + idfs +  recurrence 
        
        ttot <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + idfs +  recurrence
        
        
      }
      
    } else{ #If early metastatic
      start.early.mbc <- draw_tte(1,'lnorm',coef1=2.3, coef2=log(0.2))
      
      idfs <- ifelse(start.early.mbc<idfs,start.early.mbc,idfs)
      ttot.early <- min(ifelse(start.early.mbc<idfs,start.early.mbc,idfs),ttot.early)
      
      os.mbc <- draw_tte(1,'lnorm',coef1=0.8, coef2=log(0.2)) + start.early.mbc
      
      progression.mbc <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + start.early.mbc
      
      ttot <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + start.early.mbc
      
    }
    
    os <- min(os.mbc,os.early,nat.os.s)
    

  }) %>%  add_tte(arm="noint",
                       evts = c("start","ttot", "ttot.beva","progression.mbc", "os","idfs","ttot.early","remission","recurrence","start.early.mbc"),
                       other_inp = c("os.early","os.mbc"),                       
                       input={  #reference strategy
    start <- 0

    #Early
    
    idfs <- draw_tte(1,'lnorm',coef1=2, coef2=log(0.2),beta_tx = 1.2) 
    ttot.early <- min(draw_tte(1,'lnorm',coef1=2, coef2=log(0.2),beta_tx = 1.2),idfs)
    
    os.early <- draw_tte(1,'lnorm',coef1=3, coef2=log(0.2),beta_tx = 1.2) 
    
    #if patient has remission, check when will recurrence happen
    if (fl.remission) { 
      recurrence <- idfs +draw_tte(1,'lnorm',coef1=2, coef2=log(0.2))
      remission <- idfs
      
      #if recurrence happens before death
      if (min(os.early,nat.os.s)>recurrence) { 
        
        #Late metastatic (after finishing idfs and recurrence)
        
        os.mbc <- draw_tte(1,'lnorm',coef1=0.8, coef2=log(0.2)) + idfs  +  recurrence 
        
        progression.mbc <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + idfs +  recurrence 
        
        ttot <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + idfs +  recurrence
        
      }
      
    } else{ #If early metastatic
      start.early.mbc <- draw_tte(1,'lnorm',coef1=2.3, coef2=log(0.2))
      
      idfs <- ifelse(start.early.mbc<idfs,start.early.mbc,idfs)
      ttot.early <- min(ifelse(start.early.mbc<idfs,start.early.mbc,idfs),ttot.early)
      
      os.mbc <- draw_tte(1,'lnorm',coef1=0.8, coef2=log(0.2)) + start.early.mbc
      
      progression.mbc <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + start.early.mbc
      
      ttot <- draw_tte(1,'lnorm',coef1=0.5, coef2=log(0.2)) + start.early.mbc
      
      
    }    
    os <- min(os.mbc,os.early,nat.os.s)
    
  })


```

## Add Reaction to Those Events
Once the initial times of the events have been defined, we also need to declare how events react and affect each other. To do so, we use the `evt_react_list` object and the `add_reactevt` function. This function just needs to state which event is affected, and the actual reaction (usually setting flags to 1 or 0, or creating new/adjusting events).

There are a series of objects that can be used in this context to help with the reactions. Apart from the global objects and flags defined above, we can also use `curtime` for the current event time, `prevtime` for the time of the previous event, `cur_evtlist` for the named vector of events that is yet to happen for that patient, `arm` for the current treatment in the loop, `evt` for the current event being processed, `i` expresses the patient iteration, and `simulation` the specific simulation (relevant when the number of simulations is greater than 1). Furthermore, one can also call any other input/item that has been created before or create new ones. For example, we could even modify a cost/utility item by changing it directly, e.g. through `modify_item(list(cost.idfs.tx=500))`.

Item | What does it do 
------------ | -------------
`curtime` | Current event time (numeric)
`prevtime` | Time of the previous event (numeric)
`cur_evtlist` | Named vector of events that is yet to happen for that patient (named numeric vector)
`evt` | Current event being processed (character)
`i` | Patient being iterated (numeric)
`arm` | Intervention being iterated (character)
`simulation` | Simulation being iterated (numeric)
`sens` | Sensitivity analysis being iterated (numeric)

The functions to add/modify events and inputs use lists. Whenever several inputs/events are added or modified, it's recommended to group them within one function, as it reduces the computation cost. So rather than use two `modify_item` with a list of one element, it's better to group them into a single `modify_item` with a list of two elements. 

`new_event`allows to generate events and add them to the vector of events. It accepts more than one event. 
`modify_event` allows to modify events (e.g. delay death). When adding an event, the name of the events and the time of the events must be defined. When using `modify_event`, one must indicate which events are affected and what are the new times of the events. If the event specified does not exist or has already occurred, it will be ignored. `modify_event` with `create_if_null = TRUE` argument will also generate events if they don't exist. Note that one could potentially omit part of the modeling set in `init_event_list` and actually define new events dynamically through the reactions (we do that below for the `"ae"` event). However, this can have an impact in computation time, so if possible it's always better to use `init_event_list`.

`modify_item` and `modify_item_seq`allow to modify and add items. Elements defined within this function are not evaluated sequentially in `modify_item` (i.e. defining `modify_item(list(fl.new = 1, var1 = fl.new * 5)))` will give an error if `fl.new` was not defined outside this function), while `modify_item_seq` will do it sequentially at a slightly bigger computational cost, so both are left as choices for the user. Note that one can modify costs/utilities by using the construction `type_name_category`, where type is either "qaly" or "cost", name is the name (e.g., "default") and category is the category used (e.g., "instant"), so one could pass `cost_default_instant` and modify the cost.

The list of relevant functions to be used within `add_reactevt` are:

Function | What does it do | How to use it
------------ | ------------- | -------------
`modify_item()` | Adds & Modifies items/flags/variables for future events | `modify_item(list("fl.idfs.ontx"=0,"fl.tx.beva"=0))`
`modify_item_seq()` | Adds & Modifies items/flags/variables for future events sequentially | `modify_item_seq(list("fl.idfs.ontx"=0,"fl.tx.beva"=0))`
`new_event()` | Adds events to the vector of events for that patient | `new_event(rep(list("ae"=curtime + 0.001),5))`
`modify_event()` | Modifies (or creates, with `create_if_null = TRUE`) existing events by changing their time | `modify_event(list("os"=curtime +5, "ttot"=curtime+0.0001))`

The model will run until `curtime` is set to `Inf`, so the event that terminates the model (in this case, `os`), should modify `curtime` and set it to `Inf`.

Finally, note that there could be two different ways of accumulating continuous outcomes, backwards (i.e., in the example below, we would set q_default = util.sick at the sicker event, and modify the `q_default` value in the death event) and forwards (as in the example below). This option can be modified in the `run_sim` function using the `accum_backwards` argument, which assumes forwards by default.

```{r model_reaction}

evt_react_list <-
  add_reactevt(name_evt = "start",
               input = {}) %>%
  add_reactevt(name_evt = "ttot",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                 fl.mbcs.ontx <-  0 #Flag that patient is now off-treatment
                 
               }) %>%
  add_reactevt(name_evt = "ttot.beva",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                fl.tx.beva <- 0 #Flag that patient is now off-treatment
                 
               }) %>%
  add_reactevt(name_evt = "progression.mbc",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                 fl.mbcs.progression.mbc <- 0
                 fl.mbcs_2ndline <- 1 #Flag that patient is progressed and going in 2nd line
                 
                 new_event(list("2ndline_mbc" = curtime + draw_tte(1,'exp', log(0.08))/12))
                 
               }) %>%
  add_reactevt(name_evt = "idfs",
               input = {
                 q_default = if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- if(arm=="noint"){cost.idfs.txnoint* fl.idfs.ontx  + cost.idfs}else{(cost.idfs.tx) * fl.idfs.ontx + cost.tx.beva * fl.tx.beva + cost.idfs}
                 fl.idfs <- 0
                 
               }) %>%
  add_reactevt(name_evt = "ttot.early",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- if(arm=="noint"){cost.idfs.txnoint* fl.idfs.ontx  + cost.idfs}else{(cost.idfs.tx) * fl.idfs.ontx + cost.tx.beva * fl.tx.beva + cost.idfs}
                 fl.idfs.ontx <- 0
                 fl.tx.beva <- 0 #Flag that patient is now off-treatment
                 
                 n_ae <- rpois(1,lambda=0.25*(curtime -prevtime)) #1 AE every 4 years
                 
                 if (n_ae>0) {
                   new_event(rep(list("ae" = curtime + 0.0001),n_ae))
                 }
               }) %>%
  add_reactevt(name_evt = "remission",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.recurrence * fl.recurrence
                 fl.remission <- 1
                 
               }) %>%
  add_reactevt(name_evt = "recurrence",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.recurrence * fl.recurrence
                 fl.recurrence <- 1
                 fl.remission <- 0
                 fl.mbcs <- 1
                 fl.mbcs.progression.mbc <- 1 #ad-hoc for plot
                 
               }) %>%
  add_reactevt(name_evt = "start.early.mbc",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.recurrence * fl.recurrence
                 fl.mbcs <- 1
                 fl.mbcs.progression.mbc <- 1
                 
               }) %>%
  add_reactevt(name_evt = "2ndline_mbc",
               input = {
                 q_default = if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                 fl.mbcs_2ndline <- 0
                 
                 n_ae <- rpois(1,lambda=0.25*(curtime -prevtime)) #1 AE every 4 years
                 
                 if (n_ae>0) {
                   new_event(rep(list("ae" = curtime + 0.0001),n_ae))
                 }
               }) %>%
  add_reactevt(name_evt = "ae",
               input = {
                                  
                 q_default = if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                 c_ae <- cost.ae
                 
                 modify_event(list("os" =max(cur_evtlist[["os"]] - 0.125,curtime +0.0001) ))#each AE brings forward death by 1.5 months
               }) %>%
  add_reactevt(name_evt = "os",
               input = {
                 q_default <- if (fl.idfs==1) {
                                util.idfs.ontx * fl.idfs.ontx + (1-fl.idfs.ontx) * (1-fl.idfs.ontx) 
                              } else if (fl.idfs==0 & fl.mbcs==0) {
                                util.remission * fl.remission + fl.recurrence*util.recurrence
                              } else if (fl.mbcs==1) {
                                util.mbc.progression.mbc * fl.mbcs.progression.mbc + (1-fl.mbcs.progression.mbc)*util.mbc.pps
                              }
                 c_default <- cost.mbc.tx  * fl.mbcs.ontx + cost.mbc.progression.mbc * fl.mbcs.progression.mbc + cost.mbc.pps * (1-fl.mbcs.progression.mbc) + cost.2ndline*fl.mbcs_2ndline
                 fl.tx.beva <- 0
                 fl.mbcs.ontx <- 0
                 fl.idfs <- 0
                 fl.mbcs <- 0
                 curtime <- Inf
               }) 

```

# Costs and Utilities
Costs and utilities are introduced below. However, it's worth noting that the model is able to run without costs or utilities.

Utilities/Costs/Other outputs are defined by declaring which object belongs to utilities/costs/other outputs, and whether they need to be discounted continuously or discretely (instantaneous). These will be passed to the `run_sim` function.
## Utilities

```{r utilities}

util_ongoing <- "q_default"

```
## Costs

```{r costs}


cost_ongoing <- "c_default"

cost_instant <-  "c_ae"

```

# Model
## Model Execution
The model can be run using the function `run_sim` below. We must define the number of patients to be simulated, the number of simulations, whether we want to run a PSA or not, the strategy list, the inputs, events and reactions defined above, utilities, costs and also if we want any extra output and the level of ipd data desired to be exported.

It is worth noting that the `psa_bool` argument does not run a PSA automatically, but is rather an additional input/flag of the model that we use as a reference to determine whether we want to use a deterministic or stochastic input. As such, it could also be defined in `common_all_inputs` as the first item to be defined, and the result would be the same. However, we recommend it to be defined in `run_sim`.

Note that the distribution chosen, the number of events and the interaction between events can have a substantial impact on the running time of the model.

```{r model_run}
#Logic is: per patient, per intervention, per event, react to that event.
results <- run_sim(  
  npats=2000,                              # number of patients to be simulated
  n_sim=1,                                  # number of simulations to run
  psa_bool = FALSE,                         # use PSA or not. If n_sim > 1 and psa_bool = FALSE, then difference in outcomes is due to sampling (number of pats simulated)  
  arm_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  common_pt_inputs = common_pt_inputs,      # inputs that change within a simulation but are not affected by the intervention
  unique_pt_inputs = unique_pt_inputs,      # inputs that change within a simulation between interventions
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  cost_instant_list = cost_instant,
  input_out = c(                            # list of additional outputs (Flags, etc) that the user wants to export for each patient and event
                "os.early",
                "os.mbc",
                "nat.os.s",
                "sex_pt"
                )
          )
```
# Post-processing of Model Outputs
## Summary of Results

Once the model has been run, we can use the results and summarize them using the `summary_results_det` to print the results of the last simulation (if nsim=1, it's the deterministic case), and `summary_results_sim` to show the PSA results (with the confidence intervals). We can also use the individual patient data generated by the simulation, which we collect here to plot in the `psa_ipd` object. 

```{r post-processing_summary}


summary_results_det(results[[1]][[1]]) #will print the last simulation!

summary_results_sim(results[[1]])

psa_ipd <- bind_rows(map(results[[1]], "merged_df")) 

psa_ipd[1:10,] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```

We can also check what has been the absolute number of events per strategy.

```{r post-processing_analysis,echo=FALSE, message=FALSE}

psa_ipd %>% group_by(arm,evtname) %>% summarise(n=n()) %>% arrange(arm,-n)%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Plots

We now use the data output to plot the histograms/densities of the simulation. 

```{r post-processing_plots1, fig.width=7, fig.height=4, fig.align = 'center'}

data_plot <- results[[1]][[1]]$merged_df %>%
  filter(evtname != "start") %>%
  group_by(arm,evtname,simulation) %>%
  mutate(median = median(evttime)) %>%
  ungroup()

#Density
ggplot(data_plot) +
  geom_density(aes(fill = arm, x = evttime),
                 alpha = 0.7) +
  geom_vline(aes(xintercept=median,col=arm)) +
  facet_wrap( ~ evtname, scales = "free_y") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()



```

We can also plot the patient level QALY/costs. Note that there are several clusters in the distribution of patients according to their QALY/costs based on the pathway they took (early metastatic vs. remission and cure or recurrence). 

```{r post-processing_plots, fig.width=5, fig.height=4, message=FALSE, fig.align = 'center'}

data_qaly_cost<- psa_ipd[,.SD[1],by=.(pat_id,arm,simulation)][,.(arm,qaly=total_qalys,cost=total_costs,pat_id,simulation)]
data_qaly_cost[,ps_id:=paste(pat_id,simulation,sep="_")]


mean_data_qaly_cost <- data_qaly_cost %>% group_by(arm) %>% summarise(across(where(is.numeric),mean))

ggplot(data_qaly_cost,aes(x=qaly, y = cost, col = arm)) + 
  geom_point(alpha=0.15,shape = 21) +
  geom_point(data=mean_data_qaly_cost, aes(x=qaly, y = cost, fill = arm), shape = 21,col="black",size=3) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = .5))


```

