---
title: 'Example for a Resource Constrained Sick-Sicker-Dead model'
author: "Javier Sanchez Alvarez"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    embed-resources: true
    self-contained-math: true
vignette: >
  %\VignetteIndexEntry{example_ssd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

# Introduction
This document runs a discrete event simulation model in the context of a late oncology model to show how the functions can be used to generate a model in only a few steps. In this particular case, we document how to run a model with constrained resources.

## Main options

```{r setup}
library(WARDEN)

library(dplyr)
library(ggplot2)
library(kableExtra)
library(purrr)

```

```{r main_opt, results='hide', message=FALSE}
options(scipen = 999)
options(digits=3)
options(tibble.print_max = 50)
```

# General inputs with delayed execution
Compared to the standard sick-sicker-death model vignette, here nothing changes, except that now we add a resource which will be constrained (in a discrete manner, i.e., it accepts an integer amount of units and the resource can only be used unit-wise, like hospital beds, or doctors).

We should be careful about where we put the constrained resource, as the level at which we set it up will determine its availability to patients. Normally the intention is to have the resource being shared within an arm by patients (and cloned across arms to make sure we compare apples to apples), which means the resource should be allocated in `common_all_inputs` (as those inputs will be cloned per arm, so they will not be shared across arms unless explicitly declared so using the specific constrained objects like `resource_discrete()` or `shared_input()`). In our case, we call the resource `beds`, and we use `resource_discrete()` to set 650 beds to be shared. We also keep track of the number of beds free for the purpose of seeing it in the results.

For the sake of an example, we also create a shared input that we will be updating as a counter of patients as they go through the simulation, to showcase how inputs can be shared. Objects that are not defined in `resource_discrete()` or `shared_inputs()` will not be shared, and hard copies will be made (unless the user introduces objects which are environments, in which case by their own nature they will be modified by reference, so they will be shared across analyses, simulations, arms, and patients, so their use is not really recommended except for advanced R users).

Because of the way the constrained simulation works, it is very important to pre-define any random numbers that will be used before the simulation executes (normally at the `unique_pt_inputs` level, though depends on the case, see below for an exception and the early breast cancer model for the norm), and not to call the "r" functions (`rexp()`, `rpois()`...) in the event reactions. This is because the random state of the model will change as it goes over the model, and the nature of the loop makes it so that the random states can be mixed between patients between one simulation and another if an event TTE is altered so its position in the queue changes from one simulation to another. This is also very important if we want to compare outcomes using constrained = FALSE and constrained = TRUE (as in constrained = FALSE the random state is patient specific and can be tracked, but in constrained is much harder to do so or would violate CRAN norms).

```{r input_delayed}
#We don't need to use sensitivity_inputs here, so we don't add that object

#Put objects here that do not change on any patient or intervention loop
#We use add_item and add_item to showcase how the user can implement the inputs (either works, add_item is just faster)
common_all_inputs <-add_item(input = {
                      util.sick   <- 0.8
                      util.sicker <- 0.5
                      cost.sick   <- 3000
                      cost.sicker <- 7000
                      cost.int    <- 1000
                      coef_noint  <- log(0.2)
                      HR_int      <- 0.8
                      drc         <- 0.035 #different values than what's assumed by default
                      drq         <- 0.035
                      random_seed_sicker_i <- sample.int(100000,npats,replace = FALSE)
                      beds <- resource_discrete(650) #initialized with 650 beds 
                      beds_free <- beds$n_free() #extract current n_free
                      shared_accumulator <- shared_input(0) #initialized at 0
                      value_accum <- shared_accumulator$value() #extract value
})  


#Put objects here that do not change as we loop through treatments for a patient
common_pt_inputs <- add_item(death= max(0.0000001,rnorm(n=1, mean=12, sd=3))) 

#Put objects here that change as we loop through treatments for each patient (e.g. events can affect fl.tx, but events do not affect nat.os.s)
unique_pt_inputs <- add_item(fl.sick = 1,
                             q_default = util.sick,
                             c_default = cost.sick + if(arm=="int"){cost.int}else{0},
                             had_to_queue = 0,
                             time_in_queue = NA,
                             beds_util = 0,
                             beds_n_using = 0L)

```

# Events
## Add Initial Events
Nothing changes here relative to the standard approach.

```{r model_evts}
init_event_list <- 
  add_tte(arm=c("noint","int"), evts = c("sick","sicker","death") ,input={
    sick <- 0
    sicker <- draw_tte(1,dist="exp", coef1=coef_noint, beta_tx = ifelse(arm=="int",HR_int,1), seed = random_seed_sicker_i[i]) #this way the value would be the same if it wasn't for the HR, effectively "cloning" patients luck
    
  })

```

## Add Reaction to Those Events
We will assume that when the patients enter the sicker event, they attempt to use one of the beds. If they do so, everything goes as normal. If they fail to use one of the beds, the time to death is accelerated as they cannot get the right treatment.

The resource helpers `seize()` and `release()` read `i` and `curtime` from the calling environment automatically. `seize(beds)` returns `TRUE` (acquired), `FALSE` (queued), or `NA` (rejected when `max_queue` is set). `release(beds, resume_event = "sicker")` frees the bed for the current patient and, when a patient is queued, schedules their `"sicker"` event automatically — eliminating the manual `if(was_using & queue_size > 0) new_event(...)` pattern.

Queue wait time and the queuing flag are now tracked in C++ inside the resource object. `beds$queue_wait_time_current()` returns `NA` if the patient never queued, `0` at the moment they enter the queue, grows with elapsed time at each subsequent event while waiting, and returns the total wait time once they acquire. `beds$had_to_queue()` returns `0L` or `1L` and is a permanent per-patient flag.

```{r model_reaction}
evt_react_list <-
  add_reactevt(name_evt = "sick",
               input = {
                 value_accum   <- shared_incr(shared_accumulator, 1)
                 beds_free     <- beds$n_free()
                 beds_util     <- beds$utilization()
                 beds_n_using  <- beds$n_using()
                 time_in_queue <- beds$queue_wait_time_current()
               }) |>
  add_reactevt(name_evt = "sicker",
               input = {
                 acquired  <- seize(beds)
                 beds_free <- beds$n_free()
                 beds_util    <- beds$utilization()
                 beds_n_using <- beds$n_using()
                 if (!acquired) {
                   modify_event(c(death = max(curtime, get_event("death") * 0.8)))
                 }
                 had_to_queue  <- beds$had_to_queue()
                 time_in_queue <- beds$queue_wait_time_current()
                 q_default <- util.sicker
                 c_default <- cost.sicker + if(arm=="int"){cost.int}else{0}
                 fl.sick   <- 0
               }) |>
  add_reactevt(name_evt = "death",
               input = {
                 release(beds, resume_event = "sicker")
                 time_in_queue <- beds$queue_wait_time_current()
                 beds_free     <- beds$n_free()
                 beds_util     <- beds$utilization()
                 beds_n_using  <- beds$n_using()
                 q_default     <- 0
                 c_default     <- 0
                 curtime       <- Inf
               })

```


# Costs and Utilities


## Utilities

```{r utilities}

util_ongoing <- "q_default"

```
## Costs

```{r costs}

cost_ongoing <- "c_default"

```

# Model
## Model Execution
The model is executed in exactly the same way as before, but we just need to add `constrained = TRUE` to the arguments of `run_sim`.

```{r model_run}
#Logic is: per patient, per intervention, per event, react to that event.
results <- run_sim(  
  npats=1000,                               # number of patients to be simulated
  n_sim=1,                                  # number of simulations to run
  psa_bool = FALSE,                         # use PSA or not. If n_sim > 1 and psa_bool = FALSE, then difference in outcomes is due to sampling (number of pats simulated)  
  arm_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  common_pt_inputs = common_pt_inputs,      # inputs that change within a simulation but are not affected by the intervention
  unique_pt_inputs = unique_pt_inputs,
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  constrained = TRUE,
  ipd = 1,
  input_out = c("beds_free","beds_util","beds_n_using","had_to_queue","time_in_queue","value_accum")
)
```
# Post-processing of model outputs
## Summary of Results

```{r post-processing_summary}


summary_results_det(results[[1]][[1]]) #print first simulation

psa_ipd <- bind_rows(map(results[[1]], "merged_df")) 

psa_ipd[1:10,] |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))


```

We can also check the evolution of the free beds over time per arm. As it can be seen, 650 beds is just enough for the `int` arm to handle all the requests, however in the `noint` arm there are a few patients which must queue for some time. Note that the same thing we do to see the number of free beds could be used to understand the % of patients that were able to use a bed, etc. WARDEN is flexible in that sense, one just needs to keep track of those numbers. For example, below we can see that around 15% of the `noint` patients had to queue at a certain point. We can also track things like queuing time, etc. In this case, for these 15% of patients, the average queuing time was around 9 months. 

```{r post-processing_analysis,echo=FALSE, message=FALSE}

ggplot(psa_ipd,aes(x=evttime,y=beds_free, col = arm)) +
  geom_line()+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()

ggplot(psa_ipd |> group_by(arm,pat_id) |> summarise(had_to_queue=factor(max(had_to_queue))),aes(x=arm, fill = had_to_queue)) +
  geom_bar(position="fill")+
    scale_y_continuous(labels = scales::percent) +
  theme_bw()

ggplot(psa_ipd |> group_by(arm,pat_id) |> summarise(time_in_queue=mean(time_in_queue,na.rm=TRUE)),
       aes(x=time_in_queue, fill = arm)) +
  geom_histogram()+
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()



```

# Resource statistics

Resource utilization (`beds_util`) and units in use (`beds_n_using`) are tracked per event and exported via `input_out`. We can plot bed utilization over time just like `beds_free`:

```{r resource_stats, echo=FALSE, message=FALSE}
ggplot(psa_ipd, aes(x = evttime, y = beds_util, col = arm)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = "Bed utilization", x = "Time") +
  theme_bw()
```

The queue wait time (`time_in_queue`) and queuing flag (`had_to_queue`) are computed inside the C++ resource and read back in the reaction with `beds$queue_wait_time_current()` and `beds$had_to_queue()`. No manual bookkeeping is needed.

## Multi-resource scheduling

When a patient needs multiple resources simultaneously (e.g., a bed AND a specialist), `seize_all()` handles atomic acquisition under `"all_or_none"` policy: the patient queues only for the first unavailable resource, holding nothing else — deadlock-free by design. Below is a complete model setup pattern using two resources.

```{r multi_resource}
# common_all_inputs — both resources initialized together
common_all_inputs_multi <- add_item(input = {
  util.sick   <- 0.8; util.sicker <- 0.5
  cost.sick   <- 3000; cost.sicker <- 7000; cost.int <- 1000
  coef_noint  <- log(0.2); HR_int <- 0.8; drc <- 0.035; drq <- 0.035
  random_seed_sicker_i <- sample.int(100000, npats, replace = FALSE)
  beds        <- resource_discrete(650)
  specialists <- resource_discrete(50)
  beds_free   <- beds$n_free()
  specialists_free   <- specialists$n_free()
})

unique_pt_inputs_multi <- add_item(
  fl.sick = 1,
  q_default = util.sick,
  c_default = cost.sick + if(arm == "int") cost.int else 0,
  had_to_queue = 0,
  time_in_queue = NA,
  specialists_free = 0L,
  beds_free = 0L
)

evt_react_list_multi <-
  add_reactevt(name_evt = "sick", input = {
    beds_free        <- beds$n_free()
    specialists_free <- specialists$n_free()
    time_in_queue    <- NA
  }) |>
  add_reactevt(name_evt = "sicker", input = {
    # Patient needs BOTH a bed AND a specialist simultaneously.
    # seize_all with "all_or_none": queues only for the first unavailable resource.
    # With 650 beds vs 50 specialists, specialists saturate first, so patients
    # typically queue for specialists only, but can't use any.
    acquired <- seize_all(list(beds, specialists))
    beds_free        <- beds$n_free()
    specialists_free <- specialists$n_free()
    if (!acquired) {
      modify_event(c(death = max(curtime, get_event("death") * 0.8)))
    }
    # had_to_queue: 1 if queued for EITHER resource (check both)
    had_to_queue  <- max(beds$had_to_queue(), specialists$had_to_queue())
    # queue_wait_time_current: 0 at entry, grows while waiting, final wait on acquisition
    q_default <- util.sicker
    c_default <- cost.sicker + if(arm == "int") cost.int else 0
    fl.sick   <- 0
  }) |>
  add_reactevt(name_evt = "death", input = {
    time_in_queue <- specialists$queue_wait_time_current()
    # Release both resources; re-trigger the next patient queued for each
    release_all(list(beds, specialists), resume_event = "sicker")
    beds_free        <- beds$n_free()
    specialists_free <- specialists$n_free()
    q_default <- 0; c_default <- 0; curtime <- Inf
  })

results_multi <- run_sim(
  npats = 1000, n_sim = 1, psa_bool = FALSE,
  arm_list = c("int", "noint"),
  common_all_inputs = common_all_inputs_multi,
  common_pt_inputs = common_pt_inputs,
  unique_pt_inputs = unique_pt_inputs_multi,
  init_event_list = init_event_list,
  evt_react_list = evt_react_list_multi,
  util_ongoing_list = util_ongoing, cost_ongoing_list = cost_ongoing,
  constrained = TRUE, ipd = 1,
  input_out = c("beds_free","specialists_free","had_to_queue","time_in_queue")
)

psa_ipd <- bind_rows(map(results_multi[[1]], "merged_df")) 
ggplot(psa_ipd) +
  geom_line(aes(x=evttime,y=beds_free, col = arm))+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()

ggplot(psa_ipd) +
  geom_line(aes(x=evttime,y=specialists_free, col = arm))+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()

ggplot(psa_ipd %>%
         filter(evtname=="death") %>%
         mutate(time_in_queue = ifelse(is.na(time_in_queue),0,time_in_queue)) %>%
         group_by(pat_id, arm) %>%
         summarise(time_in_queue = max(time_in_queue,na.rm=TRUE))
       ) +
  geom_histogram(aes(x=time_in_queue, fill = arm), alpha = 0.3, col = "white", position ="identity")+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()

```

# Sensitivity analysis
## Run model with constrained = FALSE
It can be seen that with `constrained = FALSE`, the results are equivalent to the SSD standard vignette. Furthermore, if the model is run with constrained = TRUE but in such a way that the constraint does not bind (e.g., with 1000+ beds) then the results are also equivalent. If the model is run with `constrained = FALSE`, then shared inputs and resources are no longer shared. This allows to very quickly evaluate how the model outputs change when resources are constrained vs. when they are not.

```{r model_run_unc}
#Logic is: per patient, per intervention, per event, react to that event.
results <- run_sim(  
  npats=1000,                               # number of patients to be simulated
  n_sim=1,                                  # number of simulations to run
  psa_bool = FALSE,                         # use PSA or not. If n_sim > 1 and psa_bool = FALSE, then difference in outcomes is due to sampling (number of pats simulated)  
  arm_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  common_pt_inputs = common_pt_inputs,      # inputs that change within a simulation but are not affected by the intervention
  unique_pt_inputs = unique_pt_inputs,
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  constrained = FALSE,
  ipd = 1,
  input_out = c("beds_free","had_to_queue","time_in_queue")
)

summary_results_det(results[[1]][[1]]) #print first simulation

```
## Run model constrained but unbinding
If the model is run with `constrainted = TRUE` but in such a way that the constraint does not bind (e.g., with 1000+ beds) then the results are also equivalent to `constrained = FALSE`.
```{r model_run_con_unbinding}
common_all_inputs <-add_item(input = {
                      util.sick   <- 0.8
                      util.sicker <- 0.5
                      cost.sick   <- 3000
                      cost.sicker <- 7000
                      cost.int    <- 1000
                      coef_noint  <- log(0.2)
                      HR_int      <- 0.8
                      drc         <- 0.035 #different values than what's assumed by default
                      drq         <- 0.035
                      random_seed_sicker_i <- sample.int(100000,npats,replace = FALSE)
                      beds <- resource_discrete(1000)
                      beds_free <- beds$n_free()
                      shared_accumulator <- shared_input(0)
                      value_accum <- shared_accumulator$value()
})  

#Logic is: per patient, per intervention, per event, react to that event.
results <- run_sim(  
  npats=1000,                               # number of patients to be simulated
  n_sim=1,                                  # number of simulations to run
  psa_bool = FALSE,                         # use PSA or not. If n_sim > 1 and psa_bool = FALSE, then difference in outcomes is due to sampling (number of pats simulated)  
  arm_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  common_pt_inputs = common_pt_inputs,      # inputs that change within a simulation but are not affected by the intervention
  unique_pt_inputs = unique_pt_inputs,
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  constrained = TRUE,
  ipd = 1,
  input_out = c("beds_free","had_to_queue","time_in_queue")
)

summary_results_det(results[[1]][[1]]) #print first simulation

```

## Inputs
```{r dsa_inputs}
#Load some data
list_par <- list(parameter_name = list("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),
                              base_value = list(0.8,0.5,3000,7000,1000,log(0.2),0.8),
                              DSA_min = list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                              DSA_max = list(0.9,0.7,5000,9000,2000,log(0.4),0.9),
                              PSA_dist = list("rnorm","rbeta_mse","rgamma_mse","rgamma_mse","rgamma_mse","rnorm","rlnorm"),
                              a=list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)),
                              b=lapply(list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)), function(x) abs(x/5)),
                              scenario_1=list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                              scenario_2=list(0.9,0.7,5000,9000,2000,log(0.4),0.9)
                              )

common_all_inputs <-
  input_block(base      = list_par[["base_value"]],
              psa       = pick_psa(list_par[["PSA_dist"]], rep(1, length(list_par[["PSA_dist"]])),
                                   list_par[["a"]], list_par[["b"]]),
              sens      = list_par,
              names_out = list_par[["parameter_name"]],
              dsa_names = c("DSA_min", "DSA_max")) |>
  add_item(input = {
    random_seed_sicker_i = sample(1:1000, 1000, replace = FALSE)
    beds <- resource_discrete(1000)
    beds_free <- beds$n_free()
    shared_accumulator <- shared_input(0)
    value_accum <- shared_accumulator$value()
  })

```
## Model Execution
The model is executed as before, just adding `sensitivity_bool = TRUE`. The `n_sensitivity` and `sensitivity_names` arguments are auto-detected from the `input_block()` metadata — no manual configuration required.

```{r run_dsa}
results <- run_sim(
  npats=100,                               # number of patients to be simulated
  n_sim=1,                                  # number of simulations to run
  psa_bool = FALSE,                         # use PSA or not. If n_sim > 1 and psa_bool = FALSE, then difference in outcomes is due to sampling (number of pats simulated)
  arm_list = c("int", "noint"),             # intervention list
  common_all_inputs = common_all_inputs,    # inputs common that do not change within a simulation
  common_pt_inputs = common_pt_inputs,      # inputs that change within a simulation but are not affected by the intervention
  unique_pt_inputs = unique_pt_inputs,      # inputs that change within a simulation between interventions
  init_event_list = init_event_list,        # initial event list
  evt_react_list = evt_react_list,          # reaction of events
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  sensitivity_bool = TRUE,
  constrained = TRUE,
  input_out = unlist(list_par[["parameter_name"]])
)

```

## Check results
We briefly check below that indeed the engine has been changing the corresponding parameter value.
```{r dsa_check}

data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))

#Check mean value across iterations as PSA is off
data_sensitivity |> group_by(sensitivity) |> summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),mean)

```

## Model Execution, probabilistic DSA
The model is executed as before, just activating the psa_bool option

```{r run_dsa_psa}
results <- run_sim(
  npats=100,
  n_sim=6,
  psa_bool = TRUE,
  arm_list = c("int", "noint"),
  common_all_inputs = common_all_inputs,
  common_pt_inputs = common_pt_inputs,
  unique_pt_inputs = unique_pt_inputs,
  init_event_list = init_event_list,
  evt_react_list = evt_react_list,
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  sensitivity_bool = TRUE,
  constrained = TRUE,
  input_out = c(unlist(list_par[["parameter_name"]]), "beds_free")
)

```


## Check results
We briefly check below that indeed the engine has been changing the corresponding parameter value. As expected, `beds_free` is only affected by the `HR_int` and `coef_noint`, as those affect efficacy and therefore how beds are allocated.
```{r dsa_check_psa}

data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))

#Check mean value across iterations as PSA is off
data_sensitivity |> group_by(sensitivity) |> summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int", "beds_free"),mean)

```
## Model Execution, Simple PSA
The model is executed as before, just activating the `psa_bool` option and deactivating `sensitivity_bool`.

```{r run_psa}
common_all_inputs <-add_item(input = {
                      util.sick   <- 0.8
                      util.sicker <- 0.5
                      cost.sick   <- 3000
                      cost.sicker <- 7000
                      cost.int    <- 1000
                      coef_noint  <- log(0.2)
                      HR_int      <- 0.8
                      drc         <- 0.035 #different values than what's assumed by default
                      drq         <- 0.035
                      random_seed_sicker_i <- sample.int(100000,npats,replace = FALSE)
                      beds <- resource_discrete(65)
                      beds_free <- beds$n_free()
                      shared_accumulator <- shared_input(0)
                      value_accum <- shared_accumulator$value()
                      
})  


results <- run_sim(
  npats=100,
  n_sim=10,
  psa_bool = TRUE,
  arm_list = c("int", "noint"),
  common_all_inputs = common_all_inputs,
  common_pt_inputs = common_pt_inputs,
  unique_pt_inputs = unique_pt_inputs,
  init_event_list = init_event_list,
  evt_react_list = evt_react_list,
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  sensitivity_bool = FALSE,
  constrained = TRUE,
  input_out = c(unlist(list_par[["parameter_name"]]), "beds_free")
)

```

## Check results
We briefly check below that indeed the engine has been changing the corresponding parameter values, and see how the `free_beds` has been changing over time per simulation.
```{r check_psa}

data_simulation <- bind_rows(map_depth(results,2, "merged_df"))

#Check mean value across iterations as PSA is off
data_simulation |> group_by(simulation) |> summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int", "beds_free"),mean)

ggplot(data_simulation,aes(x=evttime,y=beds_free, col = factor(simulation))) +
  facet_wrap(arm~.) + 
  geom_line()+
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_bw()

```
