---
title: 'How to Use the Automatic Input Selector'
author: "Javier Sanchez Alvarez"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    embed-resources: true
    self-contained-math: true
vignette: >
  %\VignetteIndexEntry{input_selector}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)

```

```{r setup}
library("dplyr")
library("kableExtra")
library("knitr")
library("purrr")
library("MASS")
library("WARDEN")
```

```{r main_opt, results='hide', message=FALSE}
options(scipen = 999)
options(digits=3)
options(tibble.print_max = 50)
```

# Introduction
This document explains how to use the set of functions related to automatic input selector, particularly `pick_val_v`, `pick_psa`, and `create_indicators`, so that the model takes care of everything in terms of running a deterministic analysis, DSA, PSA, probabilistic DSA, or scenario analyses.

## In a Nutshell
The key function is `pick_val_v`. This function essentially hides a loop, which iterates over each of the inputs and depending on whether the parameter is a vector or single length, and whether the PSA or scenario flags are active, proceeds to select the right value. It also can consider multiple parameters being covaried at the same time in a scenario or DSA analysis.

The `pick_psa` function is a wrapper that just calls the corresponding function from its first argument, so it can be used to draw from distributions. The recommendation is to use the "r" functions like `rnorm`, `runif`, etc. The random seed is handled automatically by the model to ensure that inputs are drawn appropiately.

Finally, the `create_indicators` is a function that will generate a vector of 0 and 1 and is used for scenario analyses and DSA, as e.g., in a DSA we need to iterate over the corresponding parameters.

We can start with a simple example, see the data below, where we have a set of parameters, with some base case values, PSA parameters, DSA, scenario, and whether the parameter would be active on a PSA (vector of 0 and 1s):

```{r data_ex, echo=FALSE}

l_inputs <- list(parameter_name = list("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),
                 base_value = list(0.8,0.5,3000,7000,1000,log(0.2),0.8),
                 PSA_dist = list("rnorm","rbeta_mse","rgamma_mse","rgamma_mse","rgamma_mse","rnorm","rlnorm"),
                 a=list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)),
                 b=lapply(list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)), function(x) abs(x/5)),
                 n=as.list(rep(1,7)),
                 DSA_min = list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                 DSA_max = list(0.9,0.7,5000,9000,2000,log(0.4),0.9),
                 scenario_1=list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                 scenario_2=list(0.9,0.7,5000,9000,2000,log(0.4),0.9),
                 psa_indicators = as.list(c(rep(1,4),rep(0,3)))
                 )

as.data.frame(do.call(cbind,lapply(l_inputs,unlist)))  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## The basic setup
We'll build our case step by step. in `pick_val_v`, we need to set which values would be the base case (`base`), the PSA (`psa`), and the sensitivity values (`sens`). We also need to set whether we are currently in a PSA or not (`psa_ind`), and whether we are in a sensitivity analysis (`sens_ind`). Finally, we need to give some guidance if we are in a sensitivity analysis on which parameters to vary at this iteration (i.e., if it's iterating over the minimum range of the DSA, is it `util.sick`? is it `util.sicker`?) and to give the names of the parameters to export the named list. The simplest case is when we don't have DSA or scenario analysis. In that case the function is relatively straightforward.

We set our base case values, we set the PSA values using the parameters in our data, indicate that we are indeed in a PSA analysis and NOT in a scenario analysis (we could have both of them being `TRUE`), we indicate that if it was a scenario analysis all parameters would be included, and finally we indicate that for the PSA, the first 4 will be included and the remaining 3 will not (defaulting to the base values).

```{r no_scen, echo=TRUE}

  pick_val_v(base        = l_inputs[["base_value"]],
             psa         = pick_psa(
                 l_inputs[["PSA_dist"]],
                 l_inputs[["n"]],
                 l_inputs[["a"]],
                 l_inputs[["b"]]),
             psa_ind     = TRUE, #PSA is active
             sens_ind    = FALSE, #No scenario analysis
             names_out   = l_inputs[["parameter_name"]],
             indicator   = rep(1,7), #This is only relevant for scenario analysis, set to all 1s
             indicator_psa = l_inputs[["psa_indicators"]] #vector of 4 1s and 3 0s.
  ) 

```
See below a more complex example, now including scenario analyses/DSA, and letting the model handling more things automatically.

Note that `pick_val_v` has a few arguments set to variables: `sens_name_used`, `psa_bool`, `sensitivity_bool` and `indicators`. All except `indicators` are set by the model automatically, and are obtained from `run_sim` (from `sensitivity_names`, `psa_bool` and `sensitivity_bool`, respectively). `sens_name_used` is selected automatically by the model, so for example if `sensitivity_names = "DSA_min", "DSA_max"`, the model automatically knows which one is currently active, so it will start with `"DSA_min"`, and once it has iterated through all the relevant parameters, it will go to `"DSA_max"`. `psa_bool` and `sensitivity_bool` are boolean flags for whether the current analysis is deterministic/probabilistic and standard/scenario(or DSA). Since in this example we are not using `run_sim`, we are setting this in advance. 

`indicators` is an object that we need to set when we are setting our inputs with `add_item`, and it will be used only for scenario/DSA analyses (so when `send_ind = sensitivity_bool = TRUE`). In this case, we set it to all 1s (with length equal to the number of vectors) whenever we are not in a scenario analysis, and otherwise we use the `create_indicators` function that takes a few extra arguments.

`create_indicators` creates a vector of 0 and 1s, taking value 1 at the right index that is going to be varied. It takes a few extra objects that are given automatically by the model: `sens`, `n_sensitivity`, and `sensitivity_names`. `sensitivity_names` comes from `run_sim` as mentioned above, and `n_sensitivity` also must be declared in `run_sim`, and is the number of parameters to be varied, in this case, 7). `sens` is the index of the analysis currently being run, in this case we assume we start with the first index so it will return a vector of 1 followed by 6 0s.

Note that we do not necessarily have to use the `indicator_psa` argument of `pick_val_v`. If left as is, it will understand that all parameters would be included in the PSA. If not, we would need to provide that argument declaring which parameters are excluded or included (vector of 0 and 1s).

To recap: in this case, we artificially set some key variables beforehand, and then we execute the `pick_val_v` function. We first run it as a deterministic analysis.

```{r simple_case, echo=TRUE}

sens <- 1
n_sensitivity <- length(l_inputs[[1]])
sensitivity_names <- c("DSA_min", "DSA_max")
#Vector of length 7, a 1 followed by 6 0s, if sens was 2 it would be a c(0,1,0,0,0,0,0) vector 
create_indicators(sens,n_sensitivity*length(sensitivity_names),rep(1,length(l_inputs[[1]])))

sens_name_used <- "DSA_min"
psa_bool <- FALSE
sensitivity_bool <- FALSE

#deterministic
indicators <-  if(sensitivity_bool){ create_indicators(sens,n_sensitivity*length(sensitivity_names),rep(1,length(l_inputs[[1]])))}else{
                                rep(1,length(l_inputs[[1]]))} 

#DETERMINISTIC
as.data.frame(
  pick_val_v(base        = l_inputs[["base_value"]],
           psa         = pick_psa(
             l_inputs[["PSA_dist"]],
             l_inputs[["n"]],
             l_inputs[["a"]],
             l_inputs[["b"]]),
           sens        = l_inputs[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
           psa_ind     = psa_bool, #FALSE
           sens_ind    = sensitivity_bool, #FALSE
           indicator   = indicators, #all 1s, or a vector of 1 1 and the rest 0s.
           names_out   = l_inputs[["parameter_name"]],
           indicator_psa = l_inputs[["psa_indicators"]] 
           )
) 
```

Now it's very easy to switch to probabilistic analysis.
```{r simple_PSA, echo=TRUE}
#PSA
psa_bool <- TRUE
as.data.frame(
  pick_val_v(base        = l_inputs[["base_value"]],
           psa         = pick_psa(
             l_inputs[["PSA_dist"]],
             l_inputs[["n"]],
             l_inputs[["a"]],
             l_inputs[["b"]]),
           sens        = l_inputs[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
           psa_ind     = psa_bool, #FALSE
           sens_ind    = sensitivity_bool, #FALSE
           indicator   = indicators, #all 1s, or a vector of 1 1 and the rest 0s.
           names_out   = l_inputs[["parameter_name"]],
           indicator_psa = l_inputs[["psa_indicators"]] 
           )
)

```


And it's very easy to switch to probabilistic scenario analysis.
```{r probabilistic_DSA, echo=TRUE}
#Probabilistic DSA, first parameter being varied as sens = 1
psa_bool <- TRUE
sensitivity_bool <- TRUE
indicators <-  if(sensitivity_bool){ create_indicators(sens,n_sensitivity*length(sensitivity_names),rep(1,length(l_inputs[[1]])))}else{
                                rep(1,length(l_inputs[[1]]))} 
as.data.frame(
  pick_val_v(base        = l_inputs[["base_value"]],
           psa         = pick_psa(
             l_inputs[["PSA_dist"]],
             l_inputs[["n"]],
             l_inputs[["a"]],
             l_inputs[["b"]]),
           sens        = l_inputs[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
           psa_ind     = psa_bool, #FALSE
           sens_ind    = sensitivity_bool, #FALSE
           indicator   = indicators, #all 1s, or a vector of 1 1 and the rest 0s.
           names_out   = l_inputs[["parameter_name"]],
           indicator_psa = l_inputs[["psa_indicators"]] 
           )
) 
```

## Integrating into a model

We now use this setup in a very simple model, where we run a deterministic, probabilistic, probabilistic DSA and deterministic scenario analysis. Note that the `sens` iterator needs to be adjusted, as it just measures the total number of sensitivity iterations. If we have 2 sensitivities with 7 parameters each, sens will go from 1 to 14. As we have only 7 parameters, we need to create the `iterator_sensitivity` variable to "reset" the index back to 1, 2, 3... when it goes over 7, so that covers DSA_min and DSA_max, and for that we simply use the `sens_iterator` function.

```{r simple_model, echo=TRUE}

rm(sens, sens_name_used, sensitivity_bool, psa_bool) #remove global objects that may confuse program

i_sensitivity <-add_item(
  iterator_sensitivity = sens_iterator(sens,n_sensitivity)
  ) %>%add_item(
            indicators = if(sensitivity_bool  & sens_name_used %in% c("DSA_min", "DSA_max")){  create_indicators(iterator_sensitivity,n_sensitivity*length(sensitivity_names),rep(1,length(l_inputs[[1]]))) #only for DSA we use this approach
              }else{rep(1,length(l_inputs[[1]]))}
                              ) 

i_simple <- add_item() %>%
  add_item(
    pick_val_v(
      base = l_inputs[["base_value"]],
      psa = pick_psa(
        l_inputs[["PSA_dist"]],
        l_inputs[["n"]],
        l_inputs[["a"]],
        l_inputs[["b"]]),
      sens          = l_inputs[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
      psa_ind       = psa_bool, #FALSE
      sens_ind      = sensitivity_bool, #FALSE
      indicator     = indicators, #all 1s, or a vector of 1 1 and the rest 0s.
      names_out     = l_inputs[["parameter_name"]],
      indicator_psa = l_inputs[["psa_indicators"]] 
      )
    )

  i_arm <- add_item(q_default = util.sick,
           c_default = cost.sick + if(arm=="int"){cost.int}else{0})


init_event_list <- 
  add_tte(arm=c("noint","int"), evts = c("a1","b1") ,input={
    a1 <- 0
    b1 <- 2
  })

evt_react_list <-
  add_reactevt(name_evt = "a1",
               input = {})  %>%
  add_reactevt(name_evt = "b1",
               input = {
                 q_default = 0
                 c_default = 0 
                 curtime = Inf
               }) 

util_ongoing <- "q_default"
cost_ongoing <- "c_default"

results <- run_sim(  
  npats=5,                               
  n_sim=1,                                  
  psa_bool = FALSE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  init_event_list = init_event_list,        
  evt_react_list = evt_react_list,          
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1
)

summary_results_sim(results[[1]])  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r simple_model_psa_only, echo=TRUE}

results <- run_sim(  
  npats=5,                              
  n_sim=2,                                  
  psa_bool = TRUE,                         
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  init_event_list = init_event_list,        
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity, #this argument can also be removed since it's not used
  sensitivity_names = NULL,           #this argument can also be removed since it's not used
  sensitivity_bool = FALSE,           #this argument can also be removed since it's not used
  n_sensitivity = 1                   #this argument can also be removed since it's not used
)

summary_results_sim(results[[1]])  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r simple_model_dsa_only, echo=TRUE}
#DSA analyses, we set n_sensitivity to 7 as we need to iterate over all the parameters
results <- run_sim(  
  npats=5,                               
  n_sim=2,                                 
  psa_bool = TRUE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  init_event_list = init_event_list,       
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity,
  sensitivity_names = c("DSA_min","DSA_max"),
  sensitivity_bool = TRUE,
  n_sensitivity = length(l_inputs[[1]]), #7 parameters
  input_out = unlist(l_inputs[["parameter_name"]]) 
)

summary_results_sens(results)

data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))
data_sensitivity %>% group_by(sensitivity) %>% summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),mean)  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r simple_model_scen_only, echo=TRUE}
#Scenario analyses, we set n_sensitivity to 1 as we don't have to iterate over all parameters, each scenario is run only once
results <- run_sim(  
  npats=5,                               
  n_sim=2,                                 
  psa_bool = TRUE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  init_event_list = init_event_list,       
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity,
  sensitivity_names = c("scenario_1","scenario_2"),
  sensitivity_bool = TRUE,
  n_sensitivity = 1,
  input_out = unlist(l_inputs[["parameter_name"]])
)

summary_results_sens(results)

data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))


data_sensitivity %>% group_by(sensitivity) %>% summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),mean)  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Parameters spread across different levels
What happens when we have inputs not only at a single level (e.g., patient level), but also at simulation, patient, arm, etc levels? In that case, we can still use pick_val_v to handle things for us, but we need to be careful with the indicators, particularly in the DSA. For example, if we have 10 parameters, of which 7 are set at simulation level and 3 at patient level, pick_val_V at the simulation level will want an indicator vector of length 7 and at patient level of length 3. In the DSA, as we need to iterate at each parameter, we need to be aware of which index is being currently run, as there is the risk of varying the first parameters of both simulation and patient level at the same time. `create_indicators` can handle this for us by telling it how many parameters have gone "before" (i.e. which ones are higher in the hierarchy, so for simulation level in this case it would be 0, but for patient level it would be the previous 7 parameters set at the simulation level). See below this example.

```{r indicators_example, echo=TRUE}

#let's set the index to 8th (so it would correspond to the patient level indicator)
create_indicators(8,20,c(1,1,1,1,1,1,1),0) #all 0s

create_indicators(8,20,c(1,1,1),7) #first index is 1! because we know we are at index 8, and we have already gone through 7 iterations

```
Let's simply add a few parameters that are also now at the patient level. Note that as we are not using this parameters into the model, they will have no impact but we can still see they are really being varied. We still need to create specific indicators for each category.

```{r model_simple_split, echo=TRUE}

l_inputs_pat <- list(parameter_name = list("age","sex"),
                 base_value = list(60,1),
                 PSA_dist = list("rnorm","rbinom"),
                 a=list(60,1),
                 b=list(10,0.5),
                 n=as.list(rep(1,2)),
                 DSA_min = list(30,0),
                 DSA_max = list(80,1),
                 scenario_1=list(55,1),
                 scenario_2=list(45,0),
                 psa_indicators = as.list(rep(1,2))
                 )

i_sensitivity <- add_item(
  iterator_sensitivity = sens_iterator(sens,n_sensitivity) #resets back to 1 if it goes over n_sensitivity
  ) %>%
  add_item(
            indicators = if(sensitivity_bool  & sens_name_used %in% c("DSA_min", "DSA_max")){
              create_indicators(iterator_sensitivity, 
                                n_sensitivity*length(sensitivity_names),
                                rep(1,length(l_inputs[[1]]))) 
            }else{
                rep(1,length(l_inputs[[1]]))
              }
          ) %>%
  add_item(
            indicators_pat = if(sensitivity_bool  & sens_name_used %in% c("DSA_min", "DSA_max")){
              create_indicators(iterator_sensitivity,
                                n_sensitivity*length(sensitivity_names),
                                rep(1,length(l_inputs_pat[[1]])),
                                length(l_inputs[[1]])) 
            }else{
                rep(1,length(l_inputs_pat[[1]]))
              }
          ) 


  
i_pat <- add_item() %>%
  add_item(
    pick_val_v(
      base = l_inputs_pat[["base_value"]],
      psa = pick_psa(
        l_inputs_pat[["PSA_dist"]],
        l_inputs_pat[["n"]],
        l_inputs_pat[["a"]],
        l_inputs_pat[["b"]]),
      sens          = l_inputs_pat[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
      psa_ind       = psa_bool, #FALSE
      sens_ind      = sensitivity_bool, #TRUE
      indicator     = indicators_pat, #all 1s, or a vector of 1 1 and the rest 0s.
      names_out     = l_inputs_pat[["parameter_name"]],
      indicator_psa = l_inputs_pat[["psa_indicators"]] 
      )
    )


results <- run_sim(  
  npats=5,                               
  n_sim=2,                                 
  psa_bool = FALSE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  common_pt_inputs = i_pat,
  init_event_list = init_event_list,       
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity,
  sensitivity_names = c("DSA_min","DSA_max"),
  sensitivity_bool = TRUE,
  n_sensitivity = length(l_inputs[[1]]) + length(l_inputs_pat[[1]]), #9 parameters
  input_out = c(unlist(l_inputs[["parameter_name"]]),unlist(l_inputs_pat[["parameter_name"]]))
)

summary_results_sens(results)


data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))


data_sensitivity %>% group_by(sensitivity) %>% summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int","age","sex"),mean)  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


## Multiple parameters covaried

WARDEN also allows to have parameters being changed together in the DSA, so if the programmer believes that e.g, age and sex should take their min and max value together, that can be done by switching from a 0-1 vector to a vector which takes integer values to reflect the DSA scenario number. See below this applied, and we also assume that the utilities and costs are varied together. In this case, the indicators are simplified, as we only need 1) the correct index (provided by `iterator_sensitivity`) and 2) the dsa indicators index to understand which parameters are covaried. We also need to make sure to adjust `n_sensitivity` in `run_sim` to the new number of dsa iterations (5).

```{r model_par_covar, echo=TRUE}

l_inputs <- list(parameter_name = list("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int"),
                 base_value = list(0.8,0.5,3000,7000,1000,log(0.2),0.8),
                 PSA_dist = list("rnorm","rbeta_mse","rgamma_mse","rgamma_mse","rgamma_mse","rnorm","rlnorm"),
                 a=list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)),
                 b=lapply(list(0.8,0.5,3000,7000,1000,log(0.2),log(0.8)), function(x) abs(x/5)),
                 n=as.list(rep(1,7)),
                 DSA_min = list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                 DSA_max = list(0.9,0.7,5000,9000,2000,log(0.4),0.9),
                 scenario_1=list(0.6,0.3,1000,5000,800,log(0.1),0.5),
                 scenario_2=list(0.9,0.7,5000,9000,2000,log(0.4),0.9),
                 psa_indicators = as.list(c(rep(1,4),rep(0,3))),
                 dsa_indicators = as.list(c(1,1,2,2,2,3,4))  #covary utilities together, costs together
                 )

l_inputs_pat <- list(parameter_name = list("age","sex"),
                 base_value = list(60,1),
                 PSA_dist = list("rnorm","rbinom"),
                 a=list(60,1),
                 b=list(10,0.5),
                 n=as.list(rep(1,2)),
                 DSA_min = list(30,0),
                 DSA_max = list(80,1),
                 scenario_1=list(55,1),
                 scenario_2=list(45,0),
                 psa_indicators = as.list(rep(1,2)),
                 dsa_indicators = list(5,5) #will be 5th analysis done
                 )

i_sensitivity <- add_item(
  iterator_sensitivity = sens_iterator(sens,n_sensitivity) #resets back to 1 if it goes over n_sensitivity
  ) 


i_simple <- add_item() %>%
  add_item(
    pick_val_v(
      base = l_inputs[["base_value"]],
      psa = pick_psa(
        l_inputs[["PSA_dist"]],
        l_inputs[["n"]],
        l_inputs[["a"]],
        l_inputs[["b"]]),
      sens          = l_inputs[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
      psa_ind       = psa_bool, #FALSE
      sens_ind      = sensitivity_bool, #TRUE
      indicator     = l_inputs[["dsa_indicators"]], 
      sens_iterator = iterator_sensitivity,
      indicator_sens_binary = FALSE,
      names_out     = l_inputs[["parameter_name"]],
      indicator_psa = l_inputs[["psa_indicators"]] ,
      distributions = l_inputs[["PSA_dist"]],
      covariances   = l_inputs[["b"]]
      )
    )
  
i_pat <- add_item() %>%
  add_item(
    pick_val_v(
      base = l_inputs_pat[["base_value"]],
      psa = pick_psa(
        l_inputs_pat[["PSA_dist"]],
        l_inputs_pat[["n"]],
        l_inputs_pat[["a"]],
        l_inputs_pat[["b"]]),
      sens          = l_inputs_pat[[sens_name_used]], #e.g., sens_name_used = "DSA_min"
      psa_ind       = psa_bool, #FALSE
      sens_ind      = sensitivity_bool, #TRUE
      indicator     = l_inputs_pat[["dsa_indicators"]], 
      sens_iterator = iterator_sensitivity,
      indicator_sens_binary = FALSE,
      names_out     = l_inputs_pat[["parameter_name"]],
      indicator_psa = l_inputs_pat[["psa_indicators"]],
      distributions = l_inputs_pat[["PSA_dist"]],
      covariances   = l_inputs_pat[["b"]]
      )
    )


results <- run_sim(  
  npats=5,                               
  n_sim=2,                                 
  psa_bool = FALSE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  common_pt_inputs = i_pat,
  init_event_list = init_event_list,       
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity,
  sensitivity_names = c("DSA_min","DSA_max"),
  sensitivity_bool = TRUE,
  n_sensitivity = length(unique(l_inputs[["dsa_indicators"]])) + length(unique(l_inputs_pat[["dsa_indicators"]])), #5 parameters!!!
  input_out = c(unlist(l_inputs[["parameter_name"]]),unlist(l_inputs_pat[["parameter_name"]]))
)

summary_results_sens(results)


data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))


data_sensitivity %>% group_by(sensitivity) %>% summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int","age","sex"),mean)  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


## Parameters that are vectors

Now we are ready to handle all cases. The only approach not shown yet is when some parameters are vectors instead of length 1. In this case, the approach is the same as the previous one, as those parameters would be varied together. We add to `l_inputs_pat` a new parameter of length 2 that will be a multivariate normal. Below it can be seen for a DSA, but by now it should be straightforward to switch to probabilistic DSA, deterministic case, or standard PSA.


```{r model_vec_par, echo=TRUE}

l_inputs_pat <- list(parameter_name = list("age","sex", "v_state"),
                 base_value = list(60,1, c(10,20)),
                 PSA_dist = list("rnorm","rbinom","mvrnorm"),
                 a=list(60,1,c(10,20)),
                 b=list(10,0.5,matrix(c(2,1,4,1),2,2)),
                 n=as.list(rep(1,3)),
                 DSA_min = list(30,0, c(5,10)),
                 DSA_max = list(80,1, c(15,25)),
                 scenario_1=list(55,1, c(12,21)),
                 scenario_2=list(45,0, c(16,10)),
                 psa_indicators = list(1,1,c(1,0)), #we vary the first one but not the second one in PSA!
                 dsa_indicators = list(5,5,c(6,6))
                 )


results <- run_sim(  
  npats=5,                               
  n_sim=2,                                 
  psa_bool = FALSE,                        
  arm_list = c("int", "noint"),             
  common_all_inputs = i_simple,
  unique_pt_inputs  = i_arm,
  common_pt_inputs = i_pat,
  init_event_list = init_event_list,       
  evt_react_list = evt_react_list,         
  util_ongoing_list = util_ongoing,
  cost_ongoing_list = cost_ongoing,
  ipd = 1,
  sensitivity_inputs = i_sensitivity,
  sensitivity_names = c("DSA_min","DSA_max"),
  sensitivity_bool = TRUE,
  n_sensitivity = length(unique(l_inputs[["dsa_indicators"]])) + length(unique(l_inputs_pat[["dsa_indicators"]])), #6 parameters!
  input_out = c(unlist(l_inputs[["parameter_name"]]),unlist(l_inputs_pat[["parameter_name"]]))
)

summary_results_sens(results)


data_sensitivity <- bind_rows(map_depth(results,2, "merged_df"))

#v_state is length > 1, so it does not appear in merged_df as it would break the data. We can get it manually and print it
v_state_avg <- tibble( v_state =
    unlist(
      lapply(
        map_depth(results,2, "v_state"), function(x) {
            vecs <- list(x[[1]]$int, x[[1]]$noint, x[[2]]$int, x[[2]]$noint)
            paste(colMeans(do.call(rbind, vecs)), collapse=", ")
        }
      )
    )
  )

data_sensitivity %>% group_by(sensitivity) %>% summarise_at(c("util.sick","util.sicker","cost.sick","cost.sicker","cost.int","coef_noint","HR_int","age","sex"),mean) %>%
  bind_cols(v_state_avg)  %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
``` 

